\plaintitle

### 2017-01-11 drafts/labor/piecework_lit.tex
wordsmithing during readthrough; biggest change was merging match--girls strike paragraphs into one

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
In the first half of the \nth{19} century, George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork}, % hold foothold
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491} by the mid-19th century.
\topic{By \citeyear{hughRaynbirdTaskWork} we find
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
particularly driven toward encapsulating the manual labor of farmwork.}
He does this by contrasting two paradigms ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers an intuitive sense of piecework;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
Employers implemented piece--rates in some aspects and
time--rates in others.
The Rowan premium system, for example,
essentially paid workers
a base rate for time plus
additional pay depending on output~\cite{rowan1901premium}.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} additional compensation tied to performance,
workers were
in some senses ``task--labourers'', but
in other senses ``day--labourers''.
This was just one of several alternatives to strict time-- and piece--rate renumeration paradigms.

At the turn of the \nth{20} century, \citeauthor{taylor1914principles}
--- a mechanical engineer with a fascination in work efficiency ---
began studying and formalizing the decomposition, tracking, and management of tasks.
In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
concretizing an idea that had nebulously been forming,
and which he had been working out himself, for years~\cite{taylor1914principles}.
Scientific Management (and Fordism) thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.

@@

\topic{Since workers could now choose their own schedule and style, there arose a discussion on how best to manage pieceworkers.}
This conversation came to regard workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
piece workers would gladly work as diligently for as long as possible because
incentive--based pay rewards hard work~\cite{clark1908cotton}.

@@

\topic{Piecework's popularity in the United States and Europe fell almost as quickly as it had climbed.}
Between 1938 and 1942, the proportion of metal workers under piecework systems had climbed steeply from 11\% to 60\%~\cite{hart2005piecework}.
By 1961, that proportion dropped to 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the holdouts of piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
\citeauthor{hart2013rise} offer a number of explanations for the sudden demise of piecework.
% \ali{Should I add ``(especially in the United States but more broadly in the developed world)'' or something? Since a lot of developing nations still have sweatshops and piecework to some extent or another?}
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of piecework industries such as manufacturing and textiles to other countries; and
\item the quality of ``multidimensional'' work, which was too difficult to evaluate~\cite{hart2013rise}.
\end{inlinelist}

@@

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to four case studies to frame our analyses:
Airy's employment of human computers;
domestic and farm workers;
the ``match--girls'' strike;
and industrial and assembly-line workers.}
In introducing these cases at a high level,
we'll trace the history of piecework
while also framing the later analysis of the leading research threads we named earlier:
complexity, decomposition, and relationships.

@@

\topic{In the \nth{19} century, the calculation of celestial bodies had become a competitive field, and Airy needed to compute tables that would allow sailors to locate themselves by starlight from sea.}
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to human computers, or people who could compute basic functions.
These human computers ``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23.

The human computers captured several aspects of task decomposition that would become common. 
First, the work was designed such that it could be done independently and without collaboration. 
Second, the work was designed so that intermediate results could be quickly verified: Airy would have two workers each do the calculation, and another person compare their answers.
Third, Airy identified ways to decompose the large task into narrowly-trainable subtasks.

This practice ensured two outcomes that disfavored workers.
First, it eliminated any potential to advance professionally, as
workers' careers ended relatively quickly,
and without formal education in mathematics they struggled to find work for which their experience was meaningful.
Second, it limited workers' ability to organize
by ensuring that workers were in relatively little communication with each other.

@@

\topic{Match--makers were some of the first workers in mass manufacturing
to successfully rally for political causes.}
At the end of the \nth{19} century,
manufacturers employed women to assemble matchsticks in factories.
These women rallied first in the form of a march on parliament in 1871 to protest a proposed tax, and 
later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.

\topic{The match--girls strike of 1888 was sparked by a worker's arbitrary docking of pay, but
much deeper resentment had been simmering for years.}
Match--girls were already frustrated with
the arbitrariness of management,
poor working conditions, and
having to work with hazardous phosphoric materials.

\topic{Regardless of the reasons, the lasting impact of the match--girls strike of 1888 was profound.}
This was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

Match--girls were some of the earliest to have formed a trade union,
according to \citeauthor{booth1903life}'s account in \citeyear{booth1903life}. % ~\cite{booth1903life}.
\citeauthor{10.2307/3827491} noted that match--girls
``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
These were workers who treasured their independence, but also fiercely protected one another. % , contributing to the common good.
Their ``brashness'' for instance may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.

@@

\topic{Piecework found its way into the war effort during World War~II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a mostly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
and correspondingly assigned to that or a similar task.

>>>>

\topic{Piecework's history traces back further than most would likely expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring teenage men
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the night sky~\cite{grier2013computers}.
In the first half of the \nth{19} century, George Airy was perhaps the first to rigorously put piecework--style decomposition to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework began in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork}, % hold foothold
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491} by the mid--\nth{19} century.
\topic{By \citeyear{hughRaynbirdTaskWork} we find
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay on piecework,
particularly driven toward encapsulating the manual labor of farmwork.}
He does this by contrasting two paradigms:
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} offers a number of illustrative examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing gives us an intuitive sense of piecework;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

\topic{It's worth acknowledging that
``this distinction [between piece--rates and time--rates] was not completely clear--cut''~\cite{hart2013rise}.}
Employers adopted piece--rates in some aspects and
time--rates in others.
The Rowan premium system, for example,
essentially paid workers
a base rate for time plus
additional pay depending on output~\cite{rowan1901premium}.
As \citeauthor{rowan1901premium}'s premium system guaranteed an hourly rate
regardless of the worker's productive output
\textit{as well as} additional compensation tied to performance,
workers were
in some senses ``task--labourers'', but
in other senses ``day--labourers''.
This was just one of several alternatives to strict time-- and piece--rate remuneration paradigms.

In the late years of the \nth{19} century, \citeauthor{taylor1914principles}
--- a mechanical engineer with an interest in work efficiency ---
began studying and formalizing the decomposition, tracking, and management of tasks~\cite{taylor1896piece}.
In \citeyear{taylor1914principles} he published \citetitle{taylor1914principles},
concretizing an idea that had nebulously been forming,
and which he had been working out himself, for years~\cite{taylor1914principles}.
Scientific Management (and Fordism) thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.

@@

\topic{Since workers could now choose their own schedule and style, a discussion arose on how best to manage pieceworkers.}
This conversation came to regard workers antagonistically~\cite{roy1954efficiency}, a far cry from the earlier rhetoric on piecework, which promised that
pieceworkers would gladly work diligently and for as long as possible, as
incentive--based pay rewarded exactly, and thus aligned the goals of both managers and workers~\cite{clark1908cotton}.

@@

\topic{Piecework's popularity in the United States and Europe fell almost as quickly as it had climbed.}
Between 1938 and 1942, the proportion of metal workers under piecework systems had climbed steeply from 11\% to 60\%~\cite{hart2005piecework}.
By 1961, \citeauthor{carlson1982time} finds, the proportion dropped to 8\%~\cite{carlson1982time}.
He notes that, from 1973 to 1980, the holdouts of piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
\citeauthor{hart2013rise} offer a number of explanations for the sudden demise of piecework.
% \ali{Should I add ``(especially in the United States but more broadly in the developed world)'' or something? Since a lot of developing nations still have sweatshops and piecework to some extent or another?}
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of piecework industries such as manufacturing and textiles to other countries; and
\item the quality of ``multidimensional'' work, which was too difficult to evaluate~\cite{hart2013rise}.
\end{inlinelist}

@@

\subsection{Case studies in piecework}
\topic{Throughout the paper, we will return to four case studies to frame our analyses:
Airy's use of human computers;
domestic and farm workers;
the ``match--girls'' strike;
and industrial and assembly--line workers.}
In introducing these cases at a high level,
we'll trace the history of piecework
while also framing the later analysis of the leading research threads we named earlier:
complexity, decomposition, and relationships.

@@

\topic{In the \nth{19} century, the calculation of celestial bodies had become a competitive field, and Airy needed to compute tables that would allow sailors to locate themselves by starlight from sea.}
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to human computers, or people who could compute basic functions.
These human computers ``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to those that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.

The human computers captured several aspects of task decomposition that would become common. 
First, the work was designed such that it could be done independently and without collaboration. 
Second, the work was designed so that intermediate results could be quickly verified: Airy would have two workers each do the calculation, and another person compare their answers.
Third, Airy identified ways to decompose the large task into narrowly--trainable subtasks.

Some of Airy's policies were more controversial, for example
firing computers once they reached age 23.
This practice ensured two outcomes that disfavored workers.
First, it drastically reduced professional advancement, as
workers' careers ended quickly,
and without conventional backgrounds in mathematics
they later struggled to find work for which their experience was meaningful.
And second, it limited workers' ability to organize
by ensuring that workers were in little communication with each other,
and that they had almost no opportunity to recognize their circumstances and
to coordinate.

@@

\topic{Match--makers were some of the first workers in mass manufacturing
to successfully rally for political causes.}
At the end of the \nth{19} century,
manufacturers had begun to employ teenage women to assemble matchsticks in factories.
These women rallied first in the form of a march on parliament in 1871 to protest a proposed tax, and 
later (more famously) in what was later called ``the match--girls strike of 1888''~\cite{10.2307/3827491}.
\topic{This later strike was sparked by a worker's arbitrary docking of pay, but
much deeper resentment had been simmering for years.}
Match--girls were already frustrated with
the arbitrariness of management,
poor working conditions, and
having to work with hazardous materials such as white phosphorus, the improper handling of which
caused serious, painful, disfiguring medical conditions in the bones and ultimately death.

\topic{Regardless of what prompted it, the lasting impact of the match--girls strike of 1888 was profound.}
This was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals'': in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%~\cite{weyer1894history}.

Match--girls were some of the earliest to have formed a trade union,
according to \citeauthor{booth1903life}'s account in \citeyear{booth1903life}. % ~\cite{booth1903life}.
\citeauthor{10.2307/3827491} noted that match--girls
``\dots~pooled their resources to purchase their plumes and clothes~\dots~and expressed their solidarity through small [and major] strikes''~\cite{booth1903life}.
But they were also, as \citeauthor{10.2307/3827491} confesses, known for ``brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
These were workers who treasured their independence, but also fiercely protected one another. % , contributing to the common good.
``Brashness'' may have detracted from their public image, but almost undoubtedly contributed to their sense of solidarity,
making their propensity to act against such unfair treatment and poor conditions understandable and maybe predictable.

@@

\topic{Piecework found its way into the war effort during World War~II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a mostly female workforce that had neither
the formal training nor
years of experience that men would have had from apprenticeships.
% that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
and correspondingly assigned to that or a similar task.
